---
title: "20190917_Data_Wrangling_1_Practice"
author: "Kevin S.W."
date: "9/17/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(dplyr)
library(readxl)

```

# Data Wrangling 1 Part 1 - Data Import

## Load in a dataset

```{r Jeff_example}

# Reads in a dataset
# 2 types of paths; absolute (the full path) vs relative (only refers to the relevant files)
# getwd() pulls the absolute path until your current working directory

# instead uses a relative path. the "." means "start here"; ".." means "one directory up"; "..." means 2 up...
litters_data <- read_csv(file = "./data/FAS_litters.csv")   # uses relative path

# the "::" means package::function in package. Typically; if you only need 1 fx from a package, use the "::"
litters_data <- janitor::clean_names(litters_data)

#skimr::skim also "skims" the data

```

other arguments in read_* also has col_names. Default is TRUE. if FALSE, it will name it x1, x2...


```{r learning_assess}

# Practice loading FAS_pups.csv
pups_data <- read_csv(file = "./data/FAS_pups.csv")
pups_data <- janitor::clean_names(pups_data)

```


```{r Jeff_example2}

# Reads in a dataset
# 2 types of paths; absolute (the full path) vs relative (only refers to the relevant files)
# getwd() pulls the absolute path until your current working directory

# instead uses a relative path. the "." means "start here"; ".." means "one directory up"; "..." means 2 up...
litters_data2 <- read_csv(file = "./data/FAS_litters.csv",
                         skip = 10, col_names = FALSE)   # uses relative path

# the "::" means package::function in package. Typically; if you only need 1 fx from a package, use the "::"
litters_data2 <- janitor::clean_names(litters_data2)

```

During data import, R will try to automatically parse data and "guess" what category a column is (logic/double/etc)

by having col_types inside read_*, we can assign the types of data ourselves. example:

```{r}

litters_data3 <- read_csv(file = "./data/FAS_litters.csv",
                         col_types = cols(
                          `Group` = col_character(),
                          `Litter Number` = col_character(),  # `content` includes the space
                          `GD0 weight` = col_double(),
                          `GD18 weight` = col_double(),
                          `GD of Birth` = col_integer(),
                          `Pups born alive` = col_integer(),
                          `Pups dead @ birth` = col_integer(),
                         `Pups survive` = col_integer()
                         )
)

# the "::" means package::function in package. Typically; if you only need 1 fx from a package, use the "::"
litters_data3 <- janitor::clean_names(litters_data3)

```


# Read in excel file...

```{r Jeff_example_excel}

mlb11_data <- readxl::read_xlsx(path = "./data/mlb11.xlsx", 
                                range = "A1:D7"   # reads only the specified xls cells
                                )

write_csv(mlb11_data, path = "./data/mlb_subset.csv")

```

## Read in SAS...

```{r}

pulse_data <- haven::read_sas("./data/public_pulse_data.sas7bdat")

```


The reason why we use read "_" instead of "." is because using "." will make R "read it as a whole datapoint". With _ , it defaults it as a tibble while . defaults to dataframe. This makes R print out everything in the data if we use . while we only have a "preview" with _ . Furthermore, since R's base dataframe 



# Data Wrangling 1 Part 2 - Data Manipulation

Reading in and manipulating certain datasets in FAS dataset.

## Selecting relevant data and modifying the "visualized" columns

```{r}

# select(df, variable1 to get, variable2,...); the order matters i.e. if it's litter_number, group
# it will show litter # first.
select(litters_data, group, litter_number)  

# starts_with() allows to select any column that "starts with..."
# ends_with(), contains() does similar things
select(litters_data, litter_number, gd0_weight, starts_with("pups")) 

# using everything(); "everything else" will just pull whatever else remaining after the listed variables
select(litters_data, litter_number, everything())

# using (-) in front of the column will remove that column.
select(litters_data, -group)

# using (:) selects the "range" from:to
select(litters_data, litter_number, gd0_weight:pups_born_alive)

# renaming: newname = oldname. rename() only renames! while doing it within select() will also 
# rearrange while renaming
select(litters_data, GROUP = group, litter_number)

```


## Filtering!!

What if we need to focus only certain values?

```{r}

# similar syntax; filter(dataset, which column == specific value)
# the "==" is used universally to say/ask y is equal x. "=" is "assign x to y"
filter(litters_data, group == "Con7")

# the same == can be used for doubles
filter(litters_data, gd_of_birth == 20)

# can use different operators as well i.e. <, >...
filter(litters_data, gd_of_birth < 20)

# you can have multiple conditions as well
filter(litters_data, group == "Con7", gd_of_birth == 20)

# you can use the OR argument, which is the shift+backslash ( | ) although this is rarely used since (,) does the 
# same thing. You can use AND (&) for "intersection"
filter(litters_data, group == "Con7" | group == "Mod8")

# if you want to "get a range"
filter(litters_data, pups_born_alive >= 4, pups_born_alive <= 6)

# if you want to see if a + b is == something
filter(litters_data, gd0_weight + gd18_weight < 70)

# filters only the missing variables
filter(litters_data, is.na(gd0_weight))  # filters for na
filter(litters_data, !is.na(gd0_weight)) # filters "not na"

# instead, use drop_na(). This however, only removes the na in the specified column
drop_na(litters_data, gd0_weight)

# other signs/syntax
# pups_survive != 4 <- pups_survive are not 4
# !(pups_survive == 4) <- 
# group %in% c("Con7", "Con8") <- looks for more than 1 variables in the column

```

## Mutating data

Unlike select() or filter(), this actually modifies your dataframe

```{r}

# can use mutate to directly do operations between columns and add this new variable to the "last" column
# or you can overwrite existing data (i.e. to_lower())
mutate(litters_data, 
       wt_gain = gd18_weight - gd0_weight,
       group = str_to_lower(group),  # makes the "new" group to be all lower case
       group = str_to_upper(group),
)

```

## Arranging data

```{r}

# arranging based on arrange(df, which column); like "sorting"
arrange(litters_data, pups_born_alive)

#arranging in the reverse order
arrange(litters_data, desc(pups_born_alive))

# arranging in "multiple" orders -> order based on 1, then 2
arrange(litters_data, pups_born_alive, gd0_weight)

```


## Piping data

unlike the examples above where we do picking data, sort, filter, etc in sequence, we can use piping... `%>%`
The shortcut is cmd + shift + M

```{r}

litters_data = 
  read_csv("./data/FAS_litters.csv") %>%
  janitor::clean_names() %>% 
  select(-pups_survive) %>% 
  mutate(
    wt_gain = gd18_weight - gd0_weight,
    group = str_to_lower(group)) %>% 
  drop_na(gd0_weight)

# other uses for piping
# litters_data %>% view() # views the data immediately

# takes the gd0_weight from litters_data then takes the mean
litters_data %>% pull(gd0_weight) %>% mean()

```


The reason why you don't need to specify which df in piping, 

```{r}

# litters_data = 
 # read_csv("./data/FAS_litters.csv") %>%
 # janitor::clean_names() %>%  # if we look at the usage of janitor::clean_names(), the first argument is 
                               # clean_names(dat, ...) where dat is specifying which and in piping, this argument
                               # defaults to dat = .  where "." signifies the "data i just worked with"
 # select(-pups_survive)

```

# Data Wrangling Part 3 - 2019/09/24

## Tidying data

```{r}

library(tidyverse)

```

### Going from wide to long

Oftentimes, we need to tidy data into the "long format" in order to be able to plot it nicely

```{r}

## pivot_longer(data, range of columns, what that column name will be, what the values in the column will be)
## names_prefix = "" removes the listed prefix
pulse_data = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit",
    names_prefix = "bdi_score_",
    values_to = "bdi score"
    ) %>%  
  mutate(
    visit = recode(visit, "bl" = "00m") # recode() allows to re-assign x value to y value
  ) %>% 
    view()

```

### Separating in litters data
In litters data, the group variable combines the "intervention" variables and "day" variables (con7, con8, etc). We need to separate this
```{r}

## SEparate(data, column select, into = c(...list), separate after "x characters")
litters_data <- read_csv(file = "./data/FAS_litters.csv") %>% 
  janitor::clean_names() %>% 
  separate(col = group, into = c("dose", "day_of_tx"), 3)
  
```

### Pivoting wider
sometimes you want wider format as opposed to long-format (i.e. when you need to find differences, etc)

```{r}
analysis_result = tibble(
  group = c("treatment", "treatment", "placebo", "placebo"),
  time = c("pre", "post", "pre", "post"),
  mean = c(4, 8, 3.5, 4)
  ) %>% 
  pivot_wider(names_from = "time",
              values_from = "mean") ## pivot_wider(data, column names from, values from...)

```

