---
title: "20190917_Data_Wrangling_Practice"
author: "Kevin S.W."
date: "9/17/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)

```

# Data Wrangling Part 1 - Data Import

## Load in a dataset

```{r Jeff_example}

# Reads in a dataset
# 2 types of paths; absolute (the full path) vs relative (only refers to the relevant files)
# getwd() pulls the absolute path until your current working directory

# instead uses a relative path. the "." means "start here"; ".." means "one directory up"; "..." means 2 up...
litters_data <- read_csv(file = "./data/FAS_litters.csv")   # uses relative path

# the "::" means package::function in package. Typically; if you only need 1 fx from a package, use the "::"
litters_data <- janitor::clean_names(litters_data)

#skimr::skim also "skims" the data

```

other arguments in read_* also has col_names. Default is TRUE. if FALSE, it will name it x1, x2...


```{r learning_assess}

# Practice loading FAS_pups.csv
pups_data <- read_csv(file = "./data/FAS_pups.csv")
pups_data <- janitor::clean_names(pups_data)

```


```{r Jeff_example2}

# Reads in a dataset
# 2 types of paths; absolute (the full path) vs relative (only refers to the relevant files)
# getwd() pulls the absolute path until your current working directory

# instead uses a relative path. the "." means "start here"; ".." means "one directory up"; "..." means 2 up...
litters_data2 <- read_csv(file = "./data/FAS_litters.csv",
                         skip = 10, col_names = FALSE)   # uses relative path

# the "::" means package::function in package. Typically; if you only need 1 fx from a package, use the "::"
litters_data2 <- janitor::clean_names(litters_data2)

```

During data import, R will try to automatically parse data and "guess" what category a column is (logic/double/etc)

by having col_types inside read_*, we can assign the types of data ourselves. example:

```{r}

litters_data3 <- read_csv(file = "./data/FAS_litters.csv",
                         col_types = cols(
                          `Group` = col_character(),
                          `Litter Number` = col_character(),  # `content` includes the space
                          `GD0 weight` = col_double(),
                          `GD18 weight` = col_double(),
                          `GD of Birth` = col_integer(),
                          `Pups born alive` = col_integer(),
                          `Pups dead @ birth` = col_integer(),
                         `Pups survive` = col_integer()
                         )
)

# the "::" means package::function in package. Typically; if you only need 1 fx from a package, use the "::"
litters_data3 <- janitor::clean_names(litters_data3)

```


# Read in excel file...

```{r Jeff_example_excel}

mlb11_data <- readxl::read_xlsx(path = "./data/mlb11.xlsx", 
                                range = "A1:D7"   # reads only the specified xls cells
                                )

write_csv(mlb11_data, path = "./data/mlb_subset.csv")

```

## Read in SAS...

```{r}

pulse_data <- haven::read_sas("./data/public_pulse_data.sas7bdat")

```


The reason why we use read "_" instead of "." is because using "." will make R "read it as a whole datapoint". With _ , it defaults it as a tibble while . defaults to dataframe. This makes R print out everything in the data if we use . while we only have a "preview" with _ . Furthermore, since R's base dataframe 



# Data Wrangling Part 2 - Data Manipulation



```{r}

```

